---
title: "random_forest_stream_T"
author: "Tao"
date:  "`r Sys.Date()`"
output: html_document
---

# load the packages
```{r, echo=FALSE, results='hide'}
#install.packages("tidyverse")
library("tidyverse")
# install.packages("dataRetrieval")
library(dataRetrieval)
# install.packages("zoo") 
library(zoo)
library(raster)
library("maptools")
library(baseflow)
#library("EcoHydRology")
library("foreign")
# install.packages("moments") 
library("moments")
library("EnvStats")
library("rnoaa")
# install.packages("gradDescent") 
library("gradDescent")
#install.packages("pracma"
library("pracma")
#install.packages("nls2")
#library("nls2")
library("prism")
require("randomForest")
```

```{r}
#the mean Aug stream temp at the USGS sites
#load("meanAugT_df.Rdata")
#the the PRISM air temperature data
#load("temp_all.Rdata")
# merged air and stream T
#load("meanAugT_all")
```

#get the mean Jul and Aug stream temp at the USGS sites
```{r}
# read the list of USGS stream temperature sites in Idaho
usgs<-read.csv("wt_ID.csv")
#usgs<-read.dbf("wt_ID.dbf") #DRAIN_SQKM

#remove non-stream site
usgs<-usgs[usgs$site_no<1e+08,]
usgs$site_no<-as.character(usgs$site_no)

meanJAT_df<-data.frame(agency_cd= " ",site_no="", Date=as.Date("3333-01-01",format="%Y-%m-%d"),X_00010_00003=0,    X_00010_00003_cd= "",X_00060_00003=0,    X_00060_00003_cd= "",yr="",mo=0 , ele=0)

#for (id in usgs) {
 for (id in usgs$site_no ) {
  siteNumber <- as.character(id) 
  parameterCd <- c("00010","00060") #00060: Discharge, cubic feet per second
 ChoptankInfo <- readNWISsite(siteNumber)
 #ChoptankInfo$alt_va : Gage datum feet above NGVD29
  statCd <- c("00003") # 00003	MEAN
  startDate <- "2011-06-01"
endDate <- "2021-09-01"
meanT <-   readNWISdv(siteNumber, parameterCd, 
                                 startDate, endDate, statCd=statCd)
  if (dim(meanT )[2]>5){
meanT$yr<- format(meanT$Date, format = "%Y") 
meanT$mo<-as.numeric(format(meanT$Date, format = "%m"))
meanJAT<-meanT[meanT$mo %in% c(7,8),c("agency_cd","site_no","Date","X_00010_00003","X_00010_00003_cd","X_00060_00003","X_00060_00003_cd","yr","mo"  )]
meanJAT$ele<-ChoptankInfo$alt_va
meanJAT_df <-rbind(meanJAT_df,meanJAT)
print(Sys.time())
}
}

# clean the data
meanJAT_df<-meanJAT_df[-1,]
meanJAT_df<-meanJAT_df[!meanJAT_df$site_no=="13185000" ,]

save(meanJAT_df,file="meanQJAT_11_21.Rdata")

```

# get the PRISM air temperature data
```{r}
setwd("C:/Users/taohuang/Documents/Tao/Data/PRISM")
temp_all<-data.frame(STAID<-c("00"),yr<-c(0),RASTERVALU<-0)
colnames(temp_all)<-c("STAID","yr","RASTERVALU")
for (yr in c(2012:2020)) {
  temp<-read.dbf(paste("temp",yr,".dbf",sep=""))
  temp$yr<-yr
    temp<-temp[,c("STAID","yr","RASTERVALU")]
    temp_all<-rbind(temp_all,temp)
}
temp_all<-temp_all[-1,]
head(temp_all)
save(temp_all, file="C:/Users/taohuang/Downloads/ICL-main/temp_all.Rdata")
```

# get the PRISM air temperature data
```{r}
prism_set_dl_dir("C:/Users/taohuang/Downloads/stream_temp_final-main")
get_prism_monthlys(type = "tmean", year = 2010:2020, mon = c(7,8), keepZip = FALSE)

```




# merge stream temperature site and air temperature to get the air temp
```{r}
meanAugT_all<-merge(meanAugT_df,temp_all,by.x=c("site_no","yr"),by.y=c("STAID","yr"))
meanAugT_all<-meanAugT_all[!is.na(meanAugT_all$X_00010_00003),]
summary(meanAugT_all)
head(meanAugT_all)
save(meanAugT_all,file="C:/Users/taohuang/Downloads/ICL-main/meanAugT_all.Rdata")
```

#monthly temp
```{r}
colnames(meanAugT_all)<-c(colnames(meanAugT_all)[1:4],"Daily_Stream_T","X_00010_00003_cd", "Daily_Q", "X_00060_00003_cd" ,"mo"  ,             "ele"         ,     "Monthly_air_T"  )

meanAugT_all_mo<-meanAugT_all %>%
  group_by(site_no  ,yr) %>%
  summarise(monthly_stream_T = mean(Daily_Stream_T), monthly_stream_Q = mean(Daily_Q),ele= mean(ele), monthly_air_T=mean(Monthly_air_T) )

head(meanAugT_all_mo)
```

# split the data into ref and non-ref streams
```{r}
#g<-read.dbf("C:/Users/taohuang/Documents/Tao/Data/gagesII_9322_point_shapefile/gagesII_9322_sept30_2011.dbf")
g<-read.dbf("/Users/owner/Downloads/gagesII_9322_point_shapefile 2/gagesII_9322_sept30_2011.dbf")

g$STAID<-as.character(g$STAID)
meanAugT_all_mo2<-merge(g,meanAugT_all_mo,by.x="STAID",by.y="site_no")
meanAugT_all_mo3<-meanAugT_all_mo2[meanAugT_all_mo2$CLASS=="Ref",]
meanAugT_all_mo4<-meanAugT_all_mo2[meanAugT_all_mo2$CLASS=="Non-ref",]
```

#plot monthly air T vs. stream T
```{r}
plot(meanAugT_all_mo$monthly_air_T, meanAugT_all_mo$monthly_stream_T)
plot(meanAugT_all_mo3$monthly_air_T,meanAugT_all_mo3$monthly_stream_T)
```

# Generate training and test data
```{r}
# I splitted dataset into training and test data. The test data will be 30% of the entire dataset.
set.seed(101)
train = sample(1:nrow(meanAugT_all_mo),  nrow(meanAugT_all_mo)*0.7 )
dim(meanAugT_all_mo)
length(train)

set.seed(101)
train2 = sample(1:nrow(meanAugT_all_mo2),  nrow(meanAugT_all_mo2)*0.7 )
dim(meanAugT_all_mo2)
length(train2)

set.seed(101)
train3 = sample(1:nrow(meanAugT_all_mo3),  nrow(meanAugT_all_mo3)*0.7 )
dim(meanAugT_all_mo3)
length(train3)

set.seed(101)
train4 = sample(1:nrow(meanAugT_all_mo4),  nrow(meanAugT_all_mo4)*0.7 )
dim(meanAugT_all_mo4)
length(train4)

```

#run the random forest models
```{r}
rf.stream_T = randomForest(monthly_stream_T ~ ele +monthly_stream_Q   +monthly_air_T , data = meanAugT_all_mo, subset = train)
rf.stream_T
rf.stream_T$importance
randomForest::varImpPlot(rf.stream_T, importance=TRUE)

rf.stream_T2 = randomForest(monthly_stream_T ~ ele +monthly_stream_Q   +monthly_air_T +CLASS , data = meanAugT_all_mo2, subset = train2)
rf.stream_T2
rf.stream_T2$importance

rf.stream_T3 = randomForest(monthly_stream_T ~ ele +monthly_stream_Q   +monthly_air_T , data = meanAugT_all_mo3, subset = train3)
rf.stream_T3
rf.stream_T3$importance

rf.stream_T4 = randomForest(monthly_stream_T ~ ele +monthly_stream_Q   +monthly_air_T , data = meanAugT_all_mo4, subset = train4)
rf.stream_T4
rf.stream_T4$importance
```

# Make prediction of the all model 
```{r}

predictions <- predict(rf.stream_T, meanAugT_all_mo[-train,] )

x=seq(1,30)

plot(meanAugT_all_mo[-train,]$monthly_stream_T,predictions,xlim=c(8,22) ,ylim=c(8,22), xlab="Observed stream temperature (degC)", ylab="Predicted stream temperature (degC)")
par(new=T)
plot(x,x,type="l",xlim=c(8,22) ,ylim=c(8,22),xlab="",ylab="")
```

# Make prediction of the ref model 
```{r}
predictions3 <- predict(rf.stream_T3, meanAugT_all_mo3[-train3,] )

plot(meanAugT_all_mo3[-train3,]$monthly_stream_T,predictions3 ,xlim=c(16,26) ,ylim=c(16,26) , xlab="Observed stream temperature (degC)", ylab="Predicted stream temperature (degC)", main="Ref")

par(new=T)
plot(x,x,type="l" ,xlim=c(16,26) ,ylim=c(16,26) ,xlab="",ylab="")
```

# Make prediction of the non-ref model 
```{r}
predictions4 <- predict(rf.stream_T4, meanAugT_all_mo4[-train4,] )

plot(meanAugT_all_mo4[-train4,]$monthly_stream_T,predictions4 ,xlim=c(16,22) ,ylim=c(16,22) , xlab="Observed stream temperature (degC)", ylab="Predicted stream temperature (degC)" , main="No ref")
par(new=T)
plot(x,x,type="l" ,xlim=c(16,22) ,ylim=c(16,22) ,xlab="",ylab="")
```
